Describing the scale and composition of calls for service in Detroit

### 2021-08-24 paper.Rmd
writing

Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for service in a typical year, defined as 2019 (pre-COVID-19). Incident descriptions (e.g., investigate person) were classified into general categories, namely, community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties. Incidents were subset for those originating from the public, and thus not initiated by an officer. Recognising that frequency counts of calls for service do not necessarily reflect the amount of police resource consumed by each incident, raw counts were supplemented by the time consumed by police officers dealing with each incident type. Findings indicated that around 55% of officer shift activity time was spent on crime incidents, with the remainder of time allocated to incidents involving medical/public health (9%), the community (7%), proactive policing (5%), quality of life (14%) and traffic duties (11%). Focusing on medical/public health incidents, findings indicated that calls concentrated in particular areas of the city and during particular hours of the day.

The data used for this study covers the city of Detroit in Michigan, United States. The City of Detroit publish comparable calls for service data through their open data portal.^[Accessible via https://data.detroitmi.gov/]. The raw data includes both citizen-initiated 911 calls to request police services and officer-initiated calls spanning back to September 2016. In alignment with @ratcliffe2021policing, the data used here is subset for the year 2019 and excludes those calls initiated by an officer. Each incident has supplementary information which describes the type of incident, the latitude and longitude coordinates, and a timestamp. For each incident, the response time and time on the scene are reported. In sum, these comprise the total time officers spent on the incident.

>>>>

Using computer aided dispatch (CAD) data from the Philadelphia Police Department (PPD) the @ratcliffe2021policing study provided a breakdown of calls for service in 2019. CAD codes, used the describe calls, of which eighty unique descriptions originated from the public and required officer dispatch, were classified into six broad categories: community issues, crime, medical/public health, proactive policing, quality of life, and traffic duties. Recognising that frequency counts of calls for service do not necessarily reflect the amount of consumed police resource, raw counts were adjusted by the time committed by officers to deal with each incident type. Findings indicated that around 55% of officer shift activity time was spent on crime incidents, with the remainder of time allocated to incidents involving medical/public health (9%), the community (7%), proactive policing (5%), quality of life (14%) and traffic duties (11%). Focusing on medical/public health incidents, findings also indicated that calls concentrated in particular areas of the city and during particular hours of the day.

The data used for this study covers the city of Detroit in Michigan, United States. The City of Detroit authority publish calls for service data through their open data portal.^[Accessible via https://data.detroitmi.gov/]. The raw data includes both citizen-initiated 911 calls to request police services and officer-initiated calls spanning back to September 2016. In alignment with @ratcliffe2021policing, the data used here is subset for the year 2019 and excludes those calls initiated by an officer. For each incident, the response time and time on the scene are reported. In sum, these comprise the total time officers spent on the incident. Incidents which had a total allocated time of zero minutes (or negative values) were excluded, leaving 301,490 calls for service throughout the year. Each incident has a _calldescription_ variable which describes the nature of the call. There were 207 unique call description categories which in the interests of parsimony and ease of interpretation were recoded into 101 categories. ^[We refer readers to the corresponding GitHub repository for further details on how call descriptions were combined (xxxxxxxxxxxxxxxxxxxxxxxxxx). Readers unfamiliar with R code can review the reference table entitled 'categorisation summary'. Each call description was then categorised into the six demand type classifications used in @ratcliffe2021policing]. For each of the six demand types, an 'other' classification was generated for those incident types were consumed less than 0.2% of police time during the year. This left a total of 48 categories for the reported breakdown.

Each incident also has corresponding latitude and longitude coordinates and a call timestamp. Due to irregularities in geocoding, likely due unknown locations for some calls, 2% of incidents were excluded from the spatial descriptive analyses.
